\documentclass[10pt]{article}
\usepackage[margin=0.5in]{geometry}
\usepackage{enumitem}
\usepackage{titlesec}
\usepackage{parskip}
\usepackage{hyperref}
\usepackage{multicol}
\usepackage{fancyhdr}
\usepackage[dvipsnames]{xcolor}

% Define custom color
\definecolor{mygreen}{HTML}{004225}
\newcommand{\sectioncolor}{\color{mygreen}}

\pagestyle{empty}

% Line between sections (no spacing)
\newcommand{\sectionline}{\noindent\rule{\linewidth}{0.4pt}}

% No extra paragraph spacing
\setlength{\parskip}{0pt}
\setlength{\parindent}{0pt}

% Tight bullet formatting
\setlist[itemize]{left=1.25em, itemsep=0pt, topsep=0pt, parsep=0pt}

% Tight section formatting with color
\titleformat{\section}
{\sectioncolor\bfseries\uppercase}
{}{0pt}{}
\titlespacing{\section}{0pt}{2pt}{2pt}

\begin{document}
	
	\begin{center}
		{\color{mygreen}
			{\Huge \textbf{Scott Nguyen}} \\
			\small
			\href{mailto:scott.nguyen8901@gmail.com}{scott.nguyen8901@gmail.com} \textbar
			(515) 943-6283 \textbar
			\href{https://www.linkedin.com/in/scottnguyen8901/}{linkedin.com/in/scottnguyen8901}
		}
	\end{center}
	
	\sectionline
	
	\section*{Education}
	
	\textbf{Master of Science in Electrical Engineering \textbar{} Space Systems Engineering} \hfill \textbf{Summer 2026} \\
	University of New Mexico \hfill GPA: 3.94/4.00
	
	\textbf{Master of Science in Aerospace Engineering} \hfill \textbf{Fall 2024} \\
	University of Illinois Urbana-Champaign
	
	\textbf{Bachelor of Science in Aerospace Engineering} \hfill \textbf{Spring 2022} \\
	Iowa State University
	
	\sectionline
	
	\section*{Experience}
	
	\textbf{Youth Development Professional} \hfill \textbf{October 2024 -- January 2025} \\
	\emph{Boys \& Girls Club}
	\begin{itemize}
		\item Designed and delivered an after-school \textbf{computer literacy} curriculum for grades 3--6 (typing, file systems, internet safety).
		\item Taught \textbf{introductory coding} via Scratch and beginner Python with project-based mini-units.
		\item Used quick pre/post checks and differentiated supports to boost engagement and skill growth.
	\end{itemize}
	
	\textbf{Math Teacher (Algebra I, Algebra II, Precalculus, Calculus)} \hfill \textbf{October 2024 -- January 2025} \\
	\emph{North Star School}
	\begin{itemize}
		\item Developed \textbf{unit plans, assessments, and problem sets} aligned to course objectives; integrated Desmos/GeoGebra.
		\item Differentiated instruction and provided timely, standards-based feedback.
		\item Connected math to \textbf{real-world STEM} applications to deepen understanding.
	\end{itemize}
	
	\textbf{Research Assistant, AE 298 National Defense Education Program} \hfill \textbf{January 2023 -- May 2024} \\
	\emph{Aerospace Engineering, University of Illinois Urbana-Champaign}
	\begin{itemize}
		\item Taught an introductory \textbf{rocketry} course for underclassmen.
		\item Collected and analyzed program data to evaluate impact.
		\item Redesigned course structure to improve retention and learning outcomes.
	\end{itemize}
	
	\textbf{Research Assistant, Grants for Advancement of Teaching in Engineering (GATE)} \hfill \textbf{November 2023 -- May 2024} \\
	\emph{Mechanical Engineering, University of Illinois Urbana-Champaign}
	\begin{itemize}
		\item Built a comprehensive \textbf{final project plan} using interdisciplinary pedagogy.
		\item Researched and authored a \textbf{literature review}; contributed to a conference manuscript.
	\end{itemize}
	
	\textbf{Teaching Assistant, ASTR 122/210} \hfill \textbf{August 2023 -- May 2024} \\
	\emph{Astronomy, University of Illinois Urbana-Champaign}
	\begin{itemize}
		\item Led weekly discussions; created exam reviews for \textbf{150+} students.
		\item Produced personalized study aids and graded \textbf{150+} assignments/exams.
	\end{itemize}
	
	\textbf{Graduate Career Peer} \hfill \textbf{August 2022 -- May 2024} \\
	\emph{Engineering Career Services, University of Illinois Urbana-Champaign}
	\begin{itemize}
		\item Advised students on career development and job-search strategies.
		\item Ran mock interviews with actionable feedback to improve performance.
	\end{itemize}
	
	\textbf{Graduate Assistant, National Defense Education Program} \hfill \textbf{January 2023 -- May 2023} \\
	\emph{Aerospace Engineering, University of Illinois Urbana-Champaign}
	\begin{itemize}
		\item Built partnerships with local \textbf{K--12} schools for rocketry/avionics initiatives.
		\item Mentored educators and students; supported impact evaluation.
	\end{itemize}
	
	\textbf{Private Tutor --- SAT/ACT Test Prep \& Math/Science} \hfill \textbf{January 2020 -- May 2022} \\
	\emph{Independent}
	\begin{itemize}
		\item Delivered one-on-one \textbf{SAT/ACT} prep (diagnostics, personalized study plans, pacing \& test-taking strategies).
		\item Tracked growth with full-length practice tests; communicated progress and next steps to students/families.
	\end{itemize}
	
	\textbf{Calculus, Physics, and Differential Equations Tutor} \hfill \textbf{January 2019 -- May 2022} \\
	\emph{Academic Success Center, Iowa State University}
	\begin{itemize}
		\item Delivered tailored lessons for diverse learners in an inclusive setting.
		\item Achieved average grade improvements of \textbf{5\%--10\%} across students.
	\end{itemize}
	
\end{document}
